---
title: "p8105_hw2_yl5221"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

```{r}
library(tidyverse)
library(readxl)
library(dplyr)
```

## Problem 1

#### Describing Data

The data set has 1868 rows and 20 columns. It contains subway line, station name, route 1-11. My data cleaning first change the column types of route8-11 to character columns. Then I cleaned names of columns. next I selected the columns needed from the csv. Last but not least, there's mutation of character to logical variable.

```{r}
trans_ent = 
  read_csv(
    "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

#### 1.

There's 465 distinct stations in different lines.

```{r}
trans_ent %>% 
  select(station_name, line) %>% 
  distinct
```

#### 2. 

There are 84 stations which are ADA compliant.

```{r}
trans_ent %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

#### 3.

The proportion is 0.377.
```{r}
trans_ent %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

#### 4.
60 distinct stations serve A train. Along A train stations, 17 of them are ada compliant.
```{r}
trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```


## Problem 2
 
#### Load in and data cleaning for mr_trash_df
```{r}
mr_trash_df = readxl::read_excel(
  "./data/Trash Wheel Collection Data.xlsx", 
  sheet = "Mr. Trash Wheel",
  range = ("A2:N550")) %>%
  janitor::clean_names() %>% 
  drop_na(., dumpster) %>% 
  mutate(sports_balls = as.integer(round(sports_balls)), df = c("mr")) 
```


#### Data cleaning for p_trash_df
```{r}
p_trash_df =
  readxl::read_excel(
    "./data/Trash Wheel Collection Data.xlsx",
    sheet = "Professor Trash Wheel",
    range = "A2:M97") %>% 
  janitor::clean_names() %>% 
  drop_na(., dumpster) %>% 
  mutate(df = c("p")) 
```

#### Join two tables
Joining two tables by "dumpster", we get 547 observations of 28 variables. Sports_balls only exist in mr_trash_df. While other variables exist in both data frames. We can use the suffix ".x" and ".y" or df variable I created to seperate them. Total weight of trash collected by Professor Trash Wheel is 190.12 tons. Total number of sports balls collected by Mr. Trash Wheel in 2020 is 856.
```{r}
trash_df =
  left_join(mr_trash_df, p_trash_df, by = "dumpster") 
```

```{r}
sum(trash_df$weight_tons.y, na.rm = TRUE)
trash_df %>% 
  filter(year.x == "2020") %>% 
  summarize(sum(sports_balls))
```
## Problem 3

#### Load in and clean pols-month.csv
```{r}
pols_df =
  read_csv(
    "./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(month = recode(as.numeric(month), "1" = "Jan", "2" = "Feb", "3" = "Mar", "4" = "Apr", "5" = "May", "6" = "Jun", "7" = "Jul", "8" = "Aug", "9" = "Sep", "10" = "Oct", "11" = "Nov", "12" = "Dec")) %>% 
  mutate(., president = ifelse(prez_gop == 1, "gop", ifelse(prez_gop == 0, "dem", "other"))) %>% 
  select(-day, -prez_dem, -gov_dem)
```

#### Load in and clean snp.csv
```{r}
snp_df =
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  separate(date, into = c("month", "day", "year"), sep = "/") %>% 
  mutate(month = recode(as.numeric(month), "1" = "Jan", "2" = "Feb", "3" = "Mar", "4" = "Apr", "5" = "May", "6" = "Jun", "7" = "Jul", "8" = "Aug", "9" = "Sep", "10" = "Oct", "11" = "Nov", "12" = "Dec")) %>% 
  mutate(year = ifelse(year <= 15, as.numeric(year) + 2000, as.numeric(year) + 1900)) %>% 
  arrange(year, month) %>% 
  relocate(year, month)
```

#### Clean unemployment.csv 
```{r}
unemployment_df =
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month", 
    values_to = "data")
```


