---
title: "p8105_hw2_yl5221"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

```{r}
library(tidyverse)
library(readxl)
library(dplyr)
```

## Problem 1

#### Describing Data

The data set has 1868 rows and 20 columns. It contains subway line, station name, route 1-11. My data cleaning first change the column types of route8-11 to character columns. Then I cleaned names of columns. next I selected the columns needed from the csv. Last but not least, there's mutation of character to logical variable.

```{r}
trans_ent = 
  read_csv(
    "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

#### 1.

There's 465 distinct stations in different lines.

```{r}
trans_ent %>% 
  select(station_name, line) %>% 
  distinct
```

#### 2. 

There are 84 stations which are ADA compliant.

```{r}
trans_ent %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

#### 3.

The proportion is 0.377.
```{r}
trans_ent %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

#### 4.
60 distinct stations serve A train. Along A train stations, 17 of them are ada compliant.
```{r}
trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```


## Problem 2
 
#### Load in and data cleaning for mr_trash_df
```{r}
mr_trash_df = readxl::read_excel(
  "./data/Trash Wheel Collection Data.xlsx", 
  sheet = "Mr. Trash Wheel",
  range = ("A2:N550")) %>%
  janitor::clean_names() %>% 
  drop_na(., dumpster) %>% 
  mutate(sports_balls = as.integer(round(sports_balls)), df = c("mr")) 
```


#### Data cleaning for p_trash_df
```{r}
p_trash_df =
  readxl::read_excel(
    "./data/Trash Wheel Collection Data.xlsx",
    sheet = "Professor Trash Wheel",
    range = "A2:M97") %>% 
  janitor::clean_names() %>% 
  drop_na(., dumpster) %>% 
  mutate(df = c("p")) 
```

#### Join two data frames
Joining two tables by "dumpster", we get 547 observations of 28 variables. Sports_balls only exist in mr_trash_df. While other variables exist in both data frames. We can use the suffix ".x" and ".y" or df variable I created to seperate them. Total weight of trash collected by Professor Trash Wheel is 190.12 tons. Total number of sports balls collected by Mr. Trash Wheel in 2020 is 856.
```{r}
trash_df =
  left_join(mr_trash_df, p_trash_df, by = "dumpster") 
```

```{r}
sum(trash_df$weight_tons.y, na.rm = TRUE)
trash_df %>% 
  filter(year.x == "2020") %>% 
  summarize(sum(sports_balls))
```
## Problem 3

#### Load in and clean pols-month.csv
```{r}
pols_df =
  read_csv(
    "./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-") %>%
  mutate(year = as.numeric(year)) %>% 
  mutate(month = recode(as.numeric(month), "1" = "Jan", "2" = "Feb", "3" = "Mar", "4" = "Apr", "5" = "May", "6" = "Jun", "7" = "Jul", "8" = "Aug", "9" = "Sep", "10" = "Oct", "11" = "Nov", "12" = "Dec")) %>% 
  mutate(., president = ifelse(prez_gop == 1, "gop", ifelse(prez_gop == 0, "dem", "other"))) %>% 
  select(-day, -prez_dem, -prez_gop)
```

#### Load in and clean snp.csv
```{r}
snp_df =
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  separate(date, into = c("month", "day", "year"), sep = "/") %>% 
  mutate(month = recode(as.numeric(month), "1" = "Jan", "2" = "Feb", "3" = "Mar", "4" = "Apr", "5" = "May", "6" = "Jun", "7" = "Jul", "8" = "Aug", "9" = "Sep", "10" = "Oct", "11" = "Nov", "12" = "Dec")) %>% 
  mutate(year = ifelse(year <= 15, as.numeric(year) + 2000, as.numeric(year) + 1900)) %>% 
  arrange(year, month) %>% 
  relocate(year, month)
```

#### Clean unemployment.csv 
```{r}
unemployment_df =
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(
    Jan:Dec,
    names_to = "month", 
    values_to = "unemployment") %>% 
   janitor::clean_names()
```

#### Join three data frames
The pols_df contains 822 rows and 9 columns. The data set includes DEM and GOP parties' data and party of the president at specific month. Year ranges from 1947-2015. The snp_df has 787 rows and 4 columns. Year ranges from 1950 to 2015. The unemployment_df contains 816 rows and 3 columns. It has the unemployment rate at specific month. Year ranges from 1948 to 2015. All three df contains year and month columns, which enables us to join them together. After joining the data frames, the final_df has 822 rows and 12 columns. 
```{r}
final_df = 
  left_join(pols_df, snp_df) %>%
  left_join(., unemployment_df)
```

